---
title: "Trump’s Tweets"
author: "Elias M. Guerra"
date: "4/19/2018"
output: github_document
---

We'll start by loading the necessary packages, loading the data, and rearranging it. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)

tt.anl <- read_csv("~/Documents/R/math311/TrumpTweets_Analysis.csv")
tt.names.wide <- tt.anl %>%
  group_by(month) %>%
  summarize(Fake.News = sum(fn),
            CNN = sum(cnn),
            NBC = sum(nbc), 
            NYTimes = sum(nyt),
            Fox = sum(fox),
            Russia = sum(russia),
            Obama = sum(obama),
            Obamacare = sum(ocare),
            Hillary.Clinton = sum(hc),
            MAGA = sum(maga))
tt.names <- gather(tt.names.wide, key = "keyword", value = "freq", Fake.News:MAGA)
```

For starters we are going to look at how frequently Trump mentions these keywords we've picked out per month. 

```{r}
# Line graph
tt.names %>% # Ew...
  ggplot() + 
  geom_line(aes(x = month, y = freq, 
                color = keyword), size = .7, alpha = .7) +
  theme(axis.text.x = element_text(angle=60, hjust=1))

```

This graph is not very informative...

```{r}
# Heat map 
ggplot(tt.names) + 
  geom_tile(aes(x = month, y = keyword, fill = freq)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

From here we can see some patterns. Just before the election Trump was tweeting very frequenctly about Hillary Clinton. And it looks like it might coincide with his tweeting "Make America Great Again". Let's take a look at the relationship.


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Clinton and MAGA
tt.names %>%
  filter(keyword %in% c("Hillary.Clinton", "MAGA")) %>%
  ggplot() + 
  geom_line(aes(x = month, y = freq, 
                color = keyword, 
                group = keyword), size = .7) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
plot(Hillary.Clinton ~ MAGA, data = tt.names.wide)
lm.Hillary <- lm(Hillary.Clinton ~ MAGA, data = tt.names.wide)
abline(lm.Hillary)
summary(lm.Hillary)
plot(lm.Hillary$residuals); abline(h = 0)
```

It looks like the relationship between the two is quite significant. The residuals looks decently normal. 

```{r}
# Heat map again
tt.names %>% filter(keyword != "Hillary.Clinton", keyword != "MAGA") %>%
  ggplot() + 
  geom_tile(aes(x = month, y = keyword, fill = freq)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```


Looking at this heatmap again (without Clinton and MAGA) we can see some more relationships that stand out. Tweets about fake news became very frequent after Trump was elected. And they might coincide with his mentions of Russia. Let's test this. 

```{r}
# Fake News and Russia
tt.names %>%
  filter(keyword %in% c("Fake.News", "Russia")) %>%
  ggplot() + 
  geom_line(aes(x = month, y = freq, 
                color = keyword, 
                group = keyword), size = .7) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
plot(Fake.News ~ Russia, data = tt.names.wide)
lm.Fake <- lm(Fake.News ~ Russia, data = tt.names.wide)
abline(lm.Fake)
summary(lm.Fake) 
plot(lm.Fake$residuals); abline(h = 0)  
```

While the relationship does seem to be very significant the residuals are all over the place. 

##### References
1. The original data set of Trump's tweets come from http://www.trumptwitterarchive.com/about
