---
title: "Trump’s Tweets"
author: "Elias M. Guerra"
date: "4/19/2018"
output: github_document
---

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(rjson)
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidytext)

trump.tweets <- fromJSON(file = "~/Documents/R/math311/condensed_2018.json")
n <- length(trump.tweets) 
trump <- matrix(nrow = n, ncol = 2)
for (i in 1:n) {
  trump[i,] <- c(trump.tweets[[i]]$created_at,
                 trump.tweets[[i]]$text)
}
trump <- as.data.frame(trump)
colnames(trump) <- c("date.time", "text")
trump$date <- as.Date(substr(trump$date.time, 1, 10), format = "%a %b %d %t")
trump$time <- substr(trump$date.time, 12, 19) 
```

Let's take a peak at our dataset. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(knitr)
kable(trump[1:3,])
```

We'll start by answer the question of who Trump is talking about the most. We'll look at the most frequent mentions. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Who does Trump mention most on Twitter
mentions <- NULL
for (i in 1:nrow(trump)) {
  x <- str_split(trump$text[i], pattern = " ", simplify = T)
  mentions <- c(mentions, str_subset(x, pattern = "^@"))
}
bad <- c("!",",", ":")
for (i in 1:3) {
  mentions <- str_replace(mentions, bad[i], "")
}
most.mentions <- 
  mentions %>%
  table %>%
  data.frame %>%
  filter(Freq >= 3) %>%
  arrange(Freq)
most.mentions$. <- factor(most.mentions$., levels = most.mentions$.)
ggplot(most.mentions) + 
  geom_bar(aes(., Freq), stat = "identity") +
  ggtitle("Who does Trump mention most on Twitter?",
          subtitle = expression("no. of mentions" >= 3)) +
  coord_flip()
```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# When does trump use positive or negative language?
trump.words <- read_csv("~/Documents/R/math311/trump_twitter_words.csv")
trump.words$date <- as.Date(trump.words$date)
nrc <- get_sentiments("nrc")
trump.sentiment <- 
  trump.words %>%
  inner_join(nrc, by = "word")
trump.sentiment %>%
  group_by(sentiment) %>%
  summarize(freq = length(date)) %>%
  ggplot + geom_bar(aes(x = sentiment, y = freq), stat = "identity") +
  coord_flip() + 
  ggtitle("What emotions does Trump use?")
trump.sentiment %>%
  group_by(date, sentiment) %>%
  summarize(freq = length(date)) %>%
  filter(sentiment %in% c("positive", "negative", "trust", "fear", "anticipation", "anger")) %>%
  ggplot() + 
  geom_smooth(aes(x = date, y = freq, color = sentiment), span = .1, se = F) + 
  ggtitle("How do the emotions of Trump's tweets change over time?")
```

IDEAS: 

1. When does Trump speak about people throughout the year?

2. What words does he use to speak about them?

3. Correlate with news/Google Trends.

4. Correlate with popularity. See 538.

5. Most commonly used words over time.

6. Capitalization: Which words does he put in all caps?